Improving the Replicability of Psychological Science Through Pedagogy

### 2016-09-23 paper/paper.Rmd
meeting changes

abstract: 
    "Replications are important to science, but who will do them? One proposal is that students can conduct direct replications as part of their training. As a proof-of-concept for this idea, here we report a series of 11 pre-registered, direct replications of findings from the 2015 volume of Psychological Science, all conducted as part of a graduate-level course. Congruent with previous studies, replications typically yielded smaller effects than originals: The modal outcome was partial support for the original claim. This work provides an estimate of the probability that a motivated graduate student can reproduce a previously published result on a first attempt, however, rather than a direct assessment of the truth of particular findings. We describe the workflow and pedagogical methods that were used in the class and discuss challenges both for the adoption of this pedagogical model and for replication research more broadly."

@@

In the current article, we report the results of replication projects conducted in a single class (a graduate-level experimental methods course). As part of the required work of the course, students conducted high-power, direct replications of published articles from the 2015 volume of the journal *Psychological Science*. These studies both give evidence about the state of replicability in the field and provide insight into the challenges and promise of pedagogical replication.

@@

In the current work, we do not address these challenges; instead, our goal is more modest. We assess the probability that a graduate student can choose an article of interest and -- in a single attempt, within constraints of budget, expertise, and effort -- reproduce the findings. We consider a number of criteria for replication success, including statistical significance, effect size, a Bayesian measure of evidence [@etz2016], and a subjective assessment with respect to the original authors' interpretations. While each of these is imperfect, taken together these measures help provide a sense of how easy it is for a student to reproduce an effect to the degree that they could confidently build on it in future work. As we discuss below, these replications may have different interpretations: While some are more direct replications, others differ on theoretically-relevant dimensions. Indeed, due to differences in population, timing, or method, some of our findings might be construed as assessing the boundary conditions of a particular effect rather than replicating the original exactly. Even in these cases, however, our work contributes to an understanding of the robustness, generalizability, and expected effect sizes of previously published findings.

@@

\begin{table}[ht]
\footnotesize
\centering
\begin{tabular}{lllllr|lrcl}
  \hline
\specialcell{Citation\\(Psychological Science; 2015)} & Expt. & \specialcell{Open\\Data?} & \specialcell{Open\\Materials?} & \specialcell{On\\MTurk?} & \specialcell{N\\(orig)} & \specialcell{Power\\Standard} & \specialcell{N\\(rep)} & \specialcell{Instructor\\Fidelity} & \specialcell{Project link} \\ 
  \hline

Atir, Rosenzweig, \& Dunning & 1b & No & Yes & Yes & 202 & Other &  50 & 6.67 & https://osf.io/uegk5/ \\ 
Ko, Sadler, \& Galinksy & 2 & Yes & Some & No &  40 & Original &  40 & 4.67 & https://osf.io/g6k2u/\\ 
Lewis \& Oyserman & 4 & No & Yes & Yes & 122 & 80\% power & 128 & 6.67 & https://osf.io/rdw6e/ \\ 
Liverence \& Scholl & 1 & No & Some & No &  18 & Original* &  19 & 5.33 & https://osf.io/sfxrb/ \\ 
Proudfoot, Kay, \& Koval & 1 & No & No & Yes &  80 & 80\% power &  84 & 6.67 & https://osf.io/ne8ph/ \\ 
Scopelliti, Loewenstein, \& Vosgerau & 3 & Yes & Yes & Yes & 550 & Other & 124 & 5.67 & https://osf.io/32wsm/ \\ 
Sofer et al. & 1 & No & Yes & No &  48 & Other &  95 & 4.33 & https://osf.io/5wmhn/ \\ 
Storm \& Stone & 3 & No & No & No &  48 & Original &  61 & 4.00 & https://osf.io/9cgpq/ \\ 
Wang et al. & 2 & No & No & No & 219 & 80\% power & 397 & 4.33 & https://osf.io/ng96w/ \\ 
Xu \& Franconeri & 1a & No & No & No &  12 & Other* &  27 & 5.00 & https://osf.io/cgvzs/ \\ 
Zaval, Markowitz, \& Weber & 1 & Yes & Yes & Yes & 312 & Original & 321 & 5.00 & https://osf.io/degwa/ \\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary characteristics of original studies and our replications, along with links to project pages for individual projects. * marks projects where the number of trials was modified.} 
\label{tab:summary}
\end{table}

@@

All projects were completed as part of a graduate-level methodology class (syllabus and materials available at [http://web.stanford.edu/class/psych254/2016/]()). At the initiation of class, all students were told that they had the opportunity to contribute their individual class assignment to a group replication project. The requirements for joining the project were to conduct a pre-registered replication of a finding from the 2015 volume of *Psychological Science* and to contribute the code, data, and materials to the writeup.

@@

All study participants were recruited on AMT. Individual sample sizes are given in Table \ref{tab:summary}. Each sample was recruited independently, using the same account but a different title. All experiments were approved by the Stanford University institutional review board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

The modal replication project in our sample was judged to be a "partial" replication, meaning that some aspects of the observed findings were different from those reported by the original authors. While this conclusion is somewhat dispiriting, it is also perhaps unsurprising given the complexity of the analyses reported in many of the original papers. In addition, `r sum(d$original_test=="interaction")` of the tests were interactions, which tend to have lower statistical power [@mcclelland1993] and have fared poorly in previous replication studies [@osc2015].

Since our sample only includes 11 studies, we describe general trends here rather than attempting to conduct statistical tests (which would be dramatically under-powered after correcting for multiple comparisons). Overall, projects had a slightly larger probability of being judged successful if they were judged to be closer to the original (`r 100 * round(mean(d$replicated_TA[d$high_fidelity == TRUE]), digits=2)`\% 
vs. 
`r 100 * round(mean(d$replicated_TA[d$high_fidelity == FALSE]), digits=2)`\%). 
There were also small numerical effects of the original study being run on Mechanical Turk (`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "Yes"]), digits=2)`\% vs. `r 100 *round(mean(d$replicated_TA[d$original_on_turk == "No"]), digits=2)`\%) and having open materials (`r 100 *round(mean(d$replicated_TA[d$open_materials!="No"]), digits=2)`\% 
vs. `r 100 *round(mean(d$replicated_TA[d$open_materials=="No"]), digits=2)`\%), though we do not believe these differences are interpretable given our sample size.

@@

```{r, fig.cap="Replication effect size (A) and Bayes factor (B), plotted by the original effect size and Bayes factor, respectively. Point size shows replication N (A) and ratio of test degrees of freedom (B), color indicates subjective replication assessments by the authors. Note that the key statistic in the Sofer et al. (2015) replication was a multivariate $F$ test, hence a comparable default Bayes factor could not be computed. Additionally, the original Bayes factor for Atir et al (2015) was many orders of magnitude greater than the others, so we displayed it at the edge of the range for visibility.", fig.pos="ht", fig.height=3.6}

@@

In this paper, we reported the results of a series of 11 direct replications of previously-published experiments from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students replicate important, recent findings in a top journal. Importantly, rather than attempting to assess the truth of particular effects, our project aimed to assess the probability that a finding -- selected on grounds of feasibility and interest -- would be replicated within the constraints of a course project.

@@

Our projects had a number of limitations, including at least: 1) limited time for iterative piloting and adjustment, 2) limits on funding for larger samples, and 3) limits on domain expertise with respect to the specific effects that were chosen. But within these relatively stringent constraints, our group was able to produce a set of close replications with generally good statistical power. Thus, our work here shows what is possible for motivated students using freely-available tools and resources. Such classes could become the backbone of future collaborative replication efforts [@frank2012;@everett2015].

>>>>

abstract: 
    "Replications are important to science, but who will do them? One proposal is that students can conduct direct replications as part of their training. As a proof-of-concept for this idea, here we report a series of 11 pre-registered, direct replications of findings from the 2015 volume of Psychological Science, all conducted as part of a graduate-level course. Congruent with previous studies, replications typically yielded smaller effects than originals: The modal outcome was partial support for the original claim. This work documents the challenges facing motivated students in reproducing previously published results on a first attempt. We describe the workflow and pedagogical methods that were used in the class and discuss challenges both for the adoption of this pedagogical model and for replication research more broadly."

@@

In the current article, we report the results of replication projects conducted in a single class (a graduate-level experimental methods course). As part of the required work of the course, students conducted high-power replications of published articles from the 2015 volume of the journal *Psychological Science*. These studies both give evidence about the state of replicability in the field and provide insight into the challenges and promise of pedagogical replication.

@@

In the current work, we do not address these issues; instead, our goal is more modest. We assess the challenges facing a student in choosing an article of interest and -- in a single attempt, within constraints of budget, expertise, and effort -- reproducing the findings. We consider a number of criteria for replication success, including statistical significance, effect size, a Bayesian measure of evidence [@etz2016], and a subjective assessment with respect to the original authors' interpretations. While each of these is imperfect, taken together these measures help provide a sense of how easy it is for a student to reproduce an effect to the degree that they could confidently build on it in future work. But, as we discuss below, the results of these replications may have different interpretations: While some are more direct, others differ on theoretically-relevant dimensions. Indeed, due to differences in population, timing, or method, some of our findings might be construed as assessing the boundary conditions of a particular effect rather than replicating the original exactly. Even in these cases, however, our work contributes to an understanding of the robustness, generalizability, and expected effect sizes of previously published findings.

@@

\begin{table}[ht]
\footnotesize
\centering
\begin{tabular}{lllllr|lrcl}
  \hline
\specialcell{Citation\\(Psychological Science; 2015)} & Expt. & \specialcell{Open\\Data?} & \specialcell{Open\\Materials?} & \specialcell{On\\MTurk?} & \specialcell{N\\(orig)} & \specialcell{Power\\Standard} & \specialcell{N\\(rep)} & \specialcell{Instructor\\Fidelity} & \specialcell{Project link} \\ 
  \hline
Atir, Rosenzweig, \& Dunning & 1b & No & Yes & Yes & 202 & Other &  50 & 6.67 & https://osf.io/uegk5/ \\ 
Ko, Sadler, \& Galinksy & 2 & Yes & Some & No &  40 & Original &  40 & 4.67 & https://osf.io/g6k2u/\\ 
Lewis \& Oyserman & 4 & No & Yes & Yes & 122 & 80\% power & 128 & 6.67 & https://osf.io/rdw6e/ \\ 
Liverence \& Scholl & 1 & No & Some & No &  18 & Original* &  19 & 5.33 & https://osf.io/sfxrb/ \\ 
Proudfoot, Kay, \& Koval & 1 & No & No & Yes &  80 & 80\% power &  84 & 6.67 & https://osf.io/ne8ph/ \\ 
Scopelliti, Loewenstein, \& Vosgerau & 3 & Yes & Yes & Yes & 550 & Other & 124 & 5.67 & https://osf.io/32wsm/ \\ 
Sofer et al. & 1 & No & Yes & No & 48 & Other &  95 & 4.33 & https://osf.io/5wmhn/ \\ 
Storm \& Stone & 3 & No & No & No & 48 & Original &  61 & 4.00 & https://osf.io/9cgpq/ \\ 
Wang et al. & 2 & No & No & No & 219 & 80\% power & 397 & 4.33 & https://osf.io/ng96w/ \\ 
Xu \& Franconeri & 1a & No & No & No &  12 & Other* &  27 & 5.00 & https://osf.io/cgvzs/ \\ 
Zaval, Markowitz, \& Weber & 1 & Yes & Yes & Yes & 312 & Original & 321 & 5.00 & https://osf.io/degwa/ \\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary characteristics of original studies and our replications, along with links to project pages for individual projects. * marks projects where the number of trials was modified.} 
\label{tab:summary}
\end{table}

@@

All projects were completed as part of a graduate-level methodology class (syllabus and materials, as well as code and data for this paper, available at [https://osf.io/98ta4](https://osf.io/98ta4)). At the initiation of class, all students were told that they had the opportunity to contribute their individual class assignment to a group replication project. The requirements for joining the project were to conduct a pre-registered replication of a finding from the 2015 volume of *Psychological Science* and to contribute the code, data, and materials to the writeup.

@@

All study participants were recruited on AMT. Individual sample sizes are given in Table \ref{tab:summary}. Each sample was recruited independently, using the same account but a different title. All experiments were approved by the Stanford University Institutional Review Board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

The modal replication project in our sample was judged to be a "partial" replication, meaning that some aspects of the observed findings were different from those reported by the original authors. Since our sample only includes 11 studies, we describe general trends here rather than attempting to conduct statistical tests (which would be dramatically under-powered after correcting for multiple comparisons). Overall, projects had a slightly larger probability of being judged successful if they were judged to be closer to the original (`r 100 * round(mean(d$replicated_TA[d$high_fidelity == TRUE]), digits=2)`\% 
vs. 
`r 100 * round(mean(d$replicated_TA[d$high_fidelity == FALSE]), digits=2)`\%). 
There were also small numerical effects of the original study being run on Mechanical Turk (`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "Yes"]), digits=2)`\% vs. `r 100 *round(mean(d$replicated_TA[d$original_on_turk == "No"]), digits=2)`\%) and having open materials (`r 100 *round(mean(d$replicated_TA[d$open_materials!="No"]), digits=2)`\% 
vs. `r 100 *round(mean(d$replicated_TA[d$open_materials=="No"]), digits=2)`\%), though we do not believe these differences are interpretable given our sample size.

@@

```{r, fig.cap="Replication effect size (A) and Bayes factor (B), plotted by the original effect size and Bayes factor, respectively. Point size shows replication N (A) and ratio of test degrees of freedom (B), color indicates subjective replication assessments by the authors. Note that the key statistic in Sofer et al. (2015) replication was a multivariate $F$ test, hence a comparable default Bayes factor could not be computed. Additionally, the original Bayes factor for Atir et al. (2015) was many orders of magnitude greater than the others, so we displayed it at the edge of the range.", fig.pos="ht", fig.height=3.6}

@@

In this paper, we reported the results of a series of 11 student replications of previously-published experiments from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students replicate important, recent findings in a top journal. Importantly, rather than attempting to assess the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale [cf. @osc2015]

@@

Our projects had a number of limitations, including at least: 1) limited time for iterative piloting and adjustment, 2) limits on funding for larger samples, and 3) limits on domain expertise with respect to the specific effects that were chosen. But within these limitations, our group was able to produce a set of relatively close replications with generally good statistical power. Thus, our work here shows what is possible for motivated students using freely-available tools and resources. Such classes could become the backbone of future collaborative replication efforts [@frank2012;@everett2015].
